
Initially proposed as a model to discern information quality in entity
resolution over large datasets\cite{tal2007algebraic}, the algebraic model
for entity resolution was later refined to describe entity resolution
itself\cite{Tal11}.
While the Fellegi-Sunter model described in Subsection~\ref{subsec:fsm} is a
probabilistic model that partitions input in two data sources to compare,
this model allows as many information sources as necessary.
Another improvement is the support for transitive matching.

\subsubsection[algdesc]{Description}\label{subsubsec:algdesc}

The core tenet of this model is that entity resolution is an \textit{
equivalence relation}, in the sense defined in set theory.

\begin{defn} If $X$ is a set and $R$ is a rule that takes two elements from
$X$, $x$ and $y$, and tells us whether $x$ is in relationship $y$ in the way
denoted with $R$, then we call $R$ a \textit{binary relation} on
$X$\cite{hoffman1971linear}.
\end{defn}

We can represent $R$ as a function $R:X \times X \rightarrow \{0,1\}$ where
$R(x,y)=1$ if $x$ is in relation $R$ with $y$ and $R(x,y)=0$ otherwise.

\begin{defn}If $R$ is a binary relation on the set $X$, it is convenient to
write $xRy$ when $R(x, y) = 1$.
A binary relation R is called:

\begin{enumerate}
    \item reflexive, if $xRx, \forall x \in X$;
    \item symmetric, if $yRx \implies xRy$, $\forall x,y \in X$;
    \item transitive, if $xRz \land yRz \implies xRz$, $\forall x,y,z \in X$.
\end{enumerate}

An \textit{equivalence relation} on $X$ is a reflexive, symmetric, and
transitive binary relation on $X$\cite{hoffman1971linear}.
\end{defn}

Note that equivalence relations seem to describe almost perfectly the
matching step in entity resolution.
Let's also observe that equivalence relations work on sets for now and move
on.

An equivalence relation over a set $X$ generates an equivalence class.

\begin{defn}Suppose $R$ is an equivalence relation on the set $X$.
If $x$ is an element of $X$, let $\bar{x}$ denote the set
$\{y \in X | xRy\}$.
$\bar{x}$ is called the \textit{equivalence class} of $x$ (for the
equivalence relation $R$)\cite{hoffman1971linear}.
\end{defn}

Since $R$ is an equivalence relation, the equivalence classes for $R$ have the
following properties\cite{hoffman1971linear}:
\begin{enumerate}
    \item $xRx \in \bar{x} \implies \bar{x} \neq \emptyset$,
    \item $y\in\bar{x}\iff~x\in\bar{y}$,$~\forall x,y \in X$ because
    equivalence relations are symmetric,
    \item $\forall x,y \in X$, $\bar{x}$ is either identical with $\bar{y}$
    or they have no elements in common\cite{hoffman1971linear,Tal11}.
\end{enumerate}

The last fundamental notion from set theory that is needed to construct the
algebraic model of entity resolution is the \textit{partition} of a set.

\begin{defn}
    A \textit{partition} of a set $X$ is a collection of non-empty disjoint
    subsets of $X$ whose union is $X$\cite{halmos1960naive,Tal11}.
\end{defn}

Things come together starting with the simple observation that an element
belonging to one of the sets from a partition over $X$ is in a relation with
itself and other elements from the same set.
We say that this relation is \textit{induced} by the partition.
Conversely, from the point of view of the above relation, we say that it
\textit{generated} the partition.

When the partition in question is the set of equivalence classes over $X$,
then the induced relation is an equivalence relation.
Conversely, an equivalence relation $R$ in $X$ will generate a partition
over $X$ comprised by the equivalence classes of $X$\cite{halmos1960naive,
Tal11}.

\begin{defn}
    The equivalence relationship over a set of entity references that exists
    when two references from that set refer to the same real-world entity is
    called \textit{entity resolution}.
\end{defn}
\textcolor{green}{daca aici ai dat definitia ER in stilul modelului algebric, ar fi bine sa incluzi si la modeulul F-S o definitie pt ER in termenii F-S}



Let's denote this relation with $ER$ and state that if $xERy$ then x and y
are \textit{linked}\cite{Tal11}.
Taking our two earlier mental notes, we can now say that applying $ER$ to
individual elements of a set is akin to the \textit{matching} step in the
entity resolution process.
With each application, we now know that $ER$ creates a partition over the
initial set.
This side-effect of equivalence relationships is the \textit{clustering}
step or aspect of the entity resolution process.
The equivalence classes from the partition generated by $ER$ are also known
as \textit{clusters} in the context of entity resolution.
\textcolor{green}{similar cu obs precedenta; ar fi bine ca si pt modelul F-S sa fie prezentata aceasta corespondenta dintre model si cei 4 pasi ai procesului ER}

A last point of notice is that entity resolution tasks do not necessarily
generate a unique outcome\cite{Tal11}.
Depending on the order in which entity references are processed, runing the
same entity resolution task might generate different partitions.
The entity resolution tasks where the order in which entity references are
processed is irrelevant are called \textit{sequence neutral}\cite{Tal11}.
In this paper we are only interested in sequence neutral entity resolution
tasks.
\textcolor{green}{similar cu obs precedenta; ar fi bine ca si pt modelul F-S sa fie prezentata aceasta perspectiva legata de cardinalitatea iesirilor}

\subsubsection[algrel]{Terminology Mapping}\label{subsubsec:algrel}

In writing this paper, we used a lot of notions that are specific to the
algebraic model.
For example, attributes and entity references preserve their meaning.
In fact, it can be safely assumed that all common terms preserve their
meaning.

This paper's terminology is divergent when it comes to our use of
information sources and our attempts to generalize data structures and the
entity resolution process itself past the limits of the algebraic model.
For example, information sources contain an inordinate amount of duplicate
data, incomplete data or noise.
The algebraic model can't function under those circumstances.
Therefore, whenever we refer to the input source for the algebraic model,
we are referring to a set of entity references containing clean data.

We also should note that the algebraic model does not concern itself with
how entity references end up being part of the same set, just that they do.
Therefore using different traits (in the sense of this paper) will lead to
different outcomes.
In fact, the only place where we feel traits might be important in relation
to the algebraic model is with regard to the sequence neutrality of an
entity resolution task.
In our view, an entity resolution task's sequence neutrality can be tested
by using separate sets of traits that change the order in which items are
fed to the entity resolution task.

The result of an entity resolution task is a partition over the input set of
entity references.
This defines the output domain of our entity resolution function as:

\begin{align}
    Res &= \{C_1,~C_2,~C_3,~\ldots,~C_n\mid~C_i\in~P\}\textrm{, so that}\nonumber\\
    C_i&\neq\emptyset,\forall~1\leq~i\leq~n\textrm{,~and}\nonumber\\
    C_i\cap~C_j&\neq\emptyset,\forall~1\leq~i\neq~j\leq~n\textrm{,~and}\nonumber\\
    \bigcup_{i=1}^{n}C_i&=Res\nonumber
\end{align}

Another way to read the above definition is that ground truth and all other
entity resolution results are partitions over $Ref$.
The definition also implies that any entity reference in $Ref$ will be a
part of only one cluster.
Each equivalence class in an entity resolution result is a cluster of entity
references.
A cluster of entity references is an entity profile.

\subsubsection[algeval]{Specific Metrics}\label{subsubsec:algeval}

The quality metrics for entity resolution under the algebraic model have to
do with computing the similarity of partitions.
In our case, these metrics are applied to compare the partitions $G$ and
$Res$.
Many such metrics exist\cite{hitesh2012} and we chose to cover some of them
which either by lineage or by their popularity have peaked our interest.

The first among these is the Rand index introduced by William Rand in
1971\cite{rand1971}.
This index computes the similarity of partitions by comparing pairs of
elements formed from the elements of the clusters of each partition.
Supposing that:
\begin{itemize}
    \item $a$ is the number of pairs that are shared between clusters of $G$
    and $Res$,
    \item $b$ is the number of pairs that are only in clusters in $G$, but
    not in clusters in $Res$,
    \item $c$ is the number of pairs that are only in clusters in $Res$, but
    not in clusters in $G$,
    \item $d$ is the total number of pairs that are not shared.
\end{itemize}
then the Rand index is defined by the quantity
$\frac{a+d}{a+b+c+d}$\cite{adjrand2001}.

The minimum value of the Rand index is 0, meaning nothing matches, while its
maximum value is 1, indicating a perfect match.
The obvious use of the Rand index is to rate the similarity of two
partitions.
Its less obvious use is to consider $1 - Rand Index$ the distance between
two partitions.

The expected value of the Rand index of two random partitions does not take
a constant value (say zero)~\cite{adjrand2001}.
That lead to the introduction of the adjusted Rand index which follows the
same principles, 
\textcolor{green}{adica 0 inseamna no match si 1 perfect match? sau la ce anume te referi prin "same principles"?}
but uses a specific distribution for
randomness~\cite{adjrand1985}.
Without going into detail, we can provide a formula for the adjusted Rand
index by using the quantities we defined for the Rand index~\cite{Tal11}.
\[
    Adjusted~Rand~Index = \frac{
        a -\left(\frac{(a+b)\cdot(a+c)}{a+b+c+d}\right)
    }{
        \frac{(2\cdot a+b+c)}{2}-\left(\frac{(a+b)\cdot(a+c)}{a+b+c+d}\right)
    }
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

The third quality metric we can use for comparing partition similarity is
the Talburt-Wang index (TWI).
The merit of this metric is that it closely follows the Rand index, but is
much simpler to compute~\cite{Tal11}.
\[
    TWI = \frac{\sqrt{|G|\cdot|Ref|}}{|G \cap Ref|},
\]
where if $X$ is a set, then $|X|$ represents the numer of elements in $X$. 
The TWI uses the number of clusters that are found in both the ground
truth and in the entity resolution task's output.
It provides a good approximation of how many clusters are returned correctly
by the entity resolution task.
The Talburt-Wang index takes values between 0 and 1, with 1 indicating that
the return value of the entity resolution task perfectly matches the ground
truth.

Other similarity metrics used to compare partitions are the pairwise and
cluster precision and recall metrics.
They were first introduced in 2010 as an alternative quality
metrics to be used for evaluating entity resolution results\cite{Men10}. 
\textcolor{green}{eu as include si referintele "radacina": https://www.jstor.org/stable/pdf/1756667.pdf, https://www.cs.utexas.edu/users/ml/papers/marlin-kdd-wkshp-03.pdf}

Similarly to the Rand Index, pairwise metrics rely on extracting pairs of
items to determine how similar two set partitions are.
They differ from the Rand index because pairwise metrics operate at the
partition level and not at the partition class level.

Pairwise precision (denoted \textit{PP}) is defined as the ratio between the
number of pairs in $Res$ that are also in $G$ and the number of pairs in
$Res$.
In other words pairwise precision tells us how many pairs the entity
resolution task got right out of the pairs it returned.
\[
    PP = \frac{|Pairs(Res) \cap Pairs(G)|}{|Pairs(Res)|}
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

Pairwise recall (denoted \textit{PR}) is defined as the ratio between the
number of pairs in $Res$ that are also in $G$ and the number of pairs in $G$.
In other words, pairwise recall tells us how many pairs the entity
resolution task got right from the ideal answer.
\[
    PR = \frac{|Pairs(Res) \cap Pairs(G)|}{|Pairs(G)|}
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

The harmonic mean between pairwise precision and pairwise recall (denoted
\textit{PF1}) is called the pairwise F1 score or the pairwise comparison
measure\cite{Men10}.
\[
    PF1=\frac{2 \cdot PP \cdot PR}{PP + PR}
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

The cluster metrics are similarly, called cluster precision, cluster recall and
cluster comparison measure (or CF1).
The cluster precision (denoted \textit{CP}) is the ratio between the number
of identical clusters between the entity resolution output and the ground
truth and the number of clusters in the entity resolution output.
In other words, we want to know how many clusters the entity resolution task
got right in relation to its own output.

\[
    CP=\frac{|Clusters(Res) \cap Clusters(G)|}{|Clusters(Res)|}
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

The cluster recall (denoted \textit{CR}) is the ratio between the number of
clusters that are common between the entity resolution task's result and
the ground truth over the number of clusters in the ground truth.
In other words, this measures how many clusters did the entity resolution
task get right in relation to the ground truth.

\[
    CR = \frac{|Clusters(Res) \cap Clusters(G)|}{|Clusters(G)|}
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

Finally, the cluster comparison metric (denoted \textit{CF1}) is the
harmonic mean between the cluster precision and the cluster recall.

\[
    CF1 = \frac{2 \cdot CP \cdot CR}{CP + CR}
\]
\textcolor{green}{tb precizat domeniul de valori posibile si ce inseamna valoare buna si valoarea rea}

The pairwise and cluster metrics provide more insight into how two entity
resolution results compare structurally, which is an ability that statistical
metrics lack.
The idea of discerning how the structures of two results compare was first
suggested in Rand's paper that described his namesake index\cite{rand1971}.

The pairwise quality metrics (the pairwise precision, recall and
comparison measure as well as the Rand Index and Adjusted Rand Index) mimic
more or less closely the statistical metrics of the F-S model.
The cluster quality metrics exposed by this model provide insights that
either complement or support the pairwise metrics.
