\documentclass[a4paper,12pt]{article}
\usepackage[T1]{fontenc}
\usepackage{ragged2e}
\usepackage{array}
\usepackage{booktabs}

\begin{document}

\section{Introduction}
\paragraph{}This document follows up on the proposal to build a
framework that would allow comparing entity matching algorithms.
Such frameworks have been built before, however the current one
aims to be
\begin{itemize}
    \item more generic in terms of the data input they allow
    \item more accurate in terms of the analysis of the results
    \item friendlier for the end-user to use
\end{itemize}
\paragraph{}In order to achieve these goals, the project benefits
from understading the types of algorithms it tackles. Moreover,
every software development process can only be parallelized so much
before needing to execute certain actions in sequence. This leads
us to the conclusion that we should prioritise the entity matching
algorithms we would experiment with.
\section{Classification Criteria}
\paragraph{}In order to do an efficient prioritisation over the entity matching
algorithms, we would classify them according to a few criteria.

\subsection{Feature Types}
\paragraph{}Browsing the literature on the subject, there's a panoply of entity
matching, entity clustering, data linkage and other articles that all refer to
the same problem. We've discussed this issue at length in the project proposal
and will not detail it further here. One interesting aspect about these
algorithms is their variety with respect to the data they're able to run on.
Some algorithms work on unstructured text to match ideas, others have structured
databases wherein they need to match table records from the same or separate 
tables. However most of them support a few select data types that they're able
to perform matching on.
\paragraph{}In order to find a good first candidate for our experiments, we will
write down the data types that are supported by each entity matching algorithm,
respectively.
\paragraph{}Another way to look at this criterion is through the lens of how
well does the input data for the algorithm need to be structured. From this
point of view, we'll also note whether the entity matching algorithm works with
structured, semi-structured or unstructured data.

\subsection{Use of an External Knowledge Base}
\paragraph{}The broadness of the entity matching problem poses another question:
does the algorithm require an external database to function or not? For example,
a matching algorithm that works across alphabets would typically require using
a support knowledge base containing those alphabets. For our experiments we're
primarily looking for entity matching solutions that work by themselves in the
first stage of the project. This is mostly due to wanting to avoid extra setup
costs.
\paragraph{}For each entity matching algorithm that we're visiting, we'll note
whether it uses an external knowledge base, for what purpose and how.

\subsection{Algorithm Family}

\paragraph{}In the current context, the algorithm family simply refers
to the class of the algorithm without going into much detail. For example,
we will refer to rule-based algorithms as \textit{rule-based}, regardless of the
distance function they use. We won't go into the details of exactly what rules
are used unless we need to (to disambiguate between two algorithms, for example).

\paragraph{}An algorithm may belong to multiple classes. In such a case,
we will use the classification assigned to it by its creator. If such a
classification wasn't provided, we'll assign it the class that suits our
interests the best. For the reader this means that our classification is
subjective and should be taken with a grain of salt.

\section{Entity Matching Algorithms}
% table of entity matching algorithms with their classification according to the
% criteria written above (TBW)
For our initial version we'll start by enumerating simple algorithms that
work on top of tabular data.

\begin{center}
    \begin{tabular}{llc}
        \toprule
        Name & Category & Reference \\
        \midrule
        Fellegi-Sunter & 2616 & tabular & \\
        ACM & 2294 & tabular & \\
        Abt & 1081 & tabular & \\
        Buy & 1092 & tabular & \\
        \bottomrule
    \end{tabular}
\end{center}

\section{Conclusion}
% the best entity matching algorithm to start with + a bit of explanation
\end{document}
